---
title: "Shiny Retirement"
author: "Julie Grantier"
date: "7/13/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(ggplot2)
library(scales)
```

## Shiny Retirement App

- Though America is aging, millions of Americans have not saved enough for retirement.  Some haven't even begun to save. This simple Shiny app can start people on the right path.
- By manipulating some basic information, a user can see just how different factors can influence their retirement savings.
- Play around with it and find a way to have a Shiny Retirement!

## A Retirement Example
```{r data}
curAge <- 20
retAge <- 65
curSavings <- 10000
annualSavings <- 6000
returnRate <- 6
inflatRate <- 2
expectedIncome <- 1000
expectedExpense <- 5000
timeTilRetirement <- retAge-curAge
timeInRetirement <- 100-retAge
endAge <- 100
rateOfReturn <- ((1 + returnRate/100)/(1 + inflatRate/100))
annualNetExpense <- 12*(expectedExpense-expectedIncome)
total <- curSavings
for(i in 1:timeTilRetirement) {
      total <- rbind(total,(total[i] + annualSavings)*rateOfReturn )
}

for(j in 1:timeInRetirement) {
      if (total[j+timeTilRetirement]>0) total <- rbind(total, (total[j+timeTilRetirement]-annualNetExpense)*rateOfReturn)
      else {
            endAge <- retAge + j-1
            break
      }
}


age <- curAge:endAge
rownames(age) <- c()
rownames(total) <- c()
moneyByAge <- data.frame(cbind(age, total))
moneyByAge <- setNames(moneyByAge, c("age","total"))
```
- Let's say you are currently `r curAge` years old and plan to retire at `r retAge`.  

- You have already saved \$`r format(curSavings, scientific=FALSE, big.mark = ",")` and plan to put an additional \$`r format(annualSavings, scientific=FALSE, big.mark = ",")` in the account each year.  

- You are expecting a return of `r returnRate`% and an inflation rate of `r inflatRate`%.

- After retirement, if you expect \$`r format(expectedIncome, scientific=FALSE, big.mark = ",")` each month in pensions and \$`r format(expectedExpense, scientific=FALSE, big.mark = ",")` in expenses, your money will last until you are about `r moneyByAge[nrow(moneyByAge),1]` years old.

- See what happens to your money in the graph on the next page.

## The Plot From Shiny Retirement

```{r cars, echo = FALSE}

           p <-  ggplot(moneyByAge, aes(x=  moneyByAge$age, y= moneyByAge$total)) + 
                  geom_area( fill="blue", alpha=.2)+ 
                  geom_line()+
                  scale_y_continuous(labels = comma)+
                  labs(x = "Your Age", y = "Your Retirement Savings")
           print(p)
```

## Where Do I Find Shiny Retirement?

- The Shiny Retirement App can be accessed here: <https://jgrantier.shinyapps.io/shiny_retirement/>

- This presentation is available here: <http://rpubs.com/jgrantier/291188>

- The source code for the App (ui.R and server.R) and the Rmarkdown for this presentation can be found on GitHub: <https://github.com/jgrantier/ShinyRetirement>



